\documentclass[11pt]{article}

\usepackage[review]{acl}
\usepackage{times}
\usepackage{latexsym}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{inconsolata}

\usepackage{graphicx}
\usepackage{amsmath} 

\usepackage{multirow}  
\usepackage{graphicx}  
\usepackage{booktabs}  
\usepackage{makecell}

\newcommand{\rot}[1]{\rotatebox{60}{\makecell{#1}}}

\newcommand{\head}[2]{\makecell[c]{\textit{#1} \\ \textit{#2}}}

\newcommand{\Ins}{Instruct }
\newcommand{\Thi}{Thinking }

\newcommand*{\affmark}[1][*]{\textsuperscript{#1}}
\usepackage{fdsymbol}


\title{FinIF: Can Your Model Follow Instructions in Financial Tasks?}


\author{Anonymous EMNLP submission}

\begin{document}
\maketitle
\begin{abstract}
Large Language Models (LLMs) are deployed in finance, where tasks are embedded in multi-stage workflows.
Unlike open-domain question answering, financial workflows impose strict delivery specifications; the viability of a generated artifact depends not only on semantic quality but also on satisfying instruction-following (IF) constraints.
We observe that even frontier models struggle with these constraints, leaving some outputs insufficient as executable and auditable downstream work products.
To bridge this gap, we introduce FinIF, a framework for evaluating and improving financial instruction following through a scalable automated data construction pipeline.
Aligned with the O*NET occupational database, FinIF structures the financial task lifecycle into five continuous workflows and extends real-world materials into FinIF-Test, a 300-item benchmark with 3,134 constraints, and FinIF-Train, a 1,064-example instruction-tuning dataset.
FinIF-Test pairs deterministic Python rule checkers for 1,346 structural constraints with calibrated LLM judges for 1,788 semantic constraints.
Empirical evaluations reveal a behavioral dichotomy: GPT-5.5 achieves 93.30\% micro-average constraint satisfaction, yet only 58.67\% strict item-level pass due to failures in citation discipline, quantitative verification, and operational thresholds.
FinIF exposes a critical transition failure of LLMs from advanced analysts to reliable workflow executors, and provides scalable data resources for financial alignment.
\end{abstract}

\input{sections/1-Introduction}
\input{sections/2-Related}
\input{sections/3-FinIF}
\input{sections/4-Experiment}
\input{sections/5-Conclusion}

\bibliography{ref}
% \input{main.bbl}
\input{sections/Appendix}

\end{document}
